---
author: "Leo XIONG"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE, error = TRUE,
                      fig.height = 3)
library(tidyverse)
library(lubridate)
source("../../scripts/viridis.R")
```

## Assignment 6

#### Due Sunday, March 26, 11:59 PM CT

### Preliminaries

- Directories
    - COURSE/homework/
    - COURSE/homework/hw06/
    - COURSE/data/
    - COURSE/scripts/
- Files
  - COURSE/homework/hw06/hw06.Rmd
  - COURSE/data/exoplanets-clean-through-2022.csv
  - COURSE/data/wordle.csv
  - COURSE/scripts/viridis.R

### Aims

- Practice strings and regular expressions


### Problems

Problems 1--4 are based on a data frame with the names of built-in R colors.
  
```{r}
df_colors = tibble(color = colors()) %>% 
  mutate(index = row_number(), .before = color)
```

  1. Create a data frame with the names of all colors whose names are exactly 7 characters long. Print a summary of this data frame with the number of colors and the first and last alphabetically in the list.
  
```{r}
colors_seven = df_colors %>% 
  filter(str_length(color) == 7)
colors_seven

  
```
  
  






  2. How many of the previous color names contain no digits? Create a second data frame which eliminates the rows from your solution to the previous problem any color names which contain a digit. Print a summary of this data frame with the number of colors and the first and last alphabetically in the list.

```{r}
colors_seven_nonums = colors_seven %>% 
  filter(!str_detect(color, "[0-9]")) %>% 
  mutate(number=1:n())
colors_seven_nonums
```





  3. Find all colors which contain consecutive letters "e". Print a summary with a count of how many such colors there are and the first and last alphabetically.
  
```{r}
df_colors %>% 
  mutate(e_count = str_count(color, "e")) %>% 
  arrange(desc(e_count), color)
```



  
  
  4. Find all color names which contain any consecutive identical letters. Create a data frame with these colors.
  
- Print the first ten rows of this data frame
- Tabulate how often each letter is used in a consecutive letter pair and print out this table arranged in descending order of the counts.
- Identify the three color names which have a distinct pair of consecutive identical letters. (These are the only color names which use the given letter pair.)
  
```{r}
consecutive = df_colors %>% 
  mutate(identical = str_extract_all(color, regex("([a-z\\d])\\1"))) %>% 
  arrange(desc(identical)) 
consecutive
```
>The three colors that have a distinct pair are "Yellow", "Green" and "Deep Pink".



  
  5. Load the exoplanet data set from the file `exoplanets-clean-through-2022.csv`. Create a new data frame by selecting only the columns `id` and `star`. Find and print all rows where the name in `star` does not match the string in `star` treated as a regular expression.
  
- What seems to be the common characteristic of such star names?

After identifying this characteristic, add a column to the `planets` data frame with the name `star_regex` which, when treated as a regular expression, does match the literal star name exactly. Such a regular expression should begin with `^` and end with `$`. Verify that each star name is matched by the regular expression.
  
  
  
```{r}
planets = read_csv("../../data/exoplanets-clean-through-2022.csv")
```
  

```{r}
planets2 = planets %>% 
  select(id, star) %>% 
  filter(!str_detect(star, star)) %>% 
  mutate(star_regex = regex(str_replace(star, "\\+", "^\\+$")))
print(planets2)
```
>The most common characteristic of the names are special characters in the star anames





  6. Most planets have a name equal to the star name followed by a space and a single lower case letter. Identify all planets for which this pattern is not true.
  
- Print the columns `id`, `planet`, `star`, and `year` for all planets in `planets` which do not fit this pattern.
  
- How many of the 5240 exoplanets do not follow the typical convention?

  
```{r}
nomatch = planets %>% 
  select(id, planet, star, year) %>% 
  filter(!str_sub(planet,-3,-3) == str_sub(star,-1,-1))
nomatch
```
>There are 45 rows where that do not follow the typical convention

  
  



  7. For the planets whose names **do match** the pattern from the previous problem, create a table which counts the frequency of the last lower case letters in the planet name. Display this data with an effective graph.
  
  
  
```{r}
match = planets %>% 
  select(id, planet, star, year) %>% 
  filter(str_sub(planet,-3,-3) == str_sub(star,-1,-1)) %>% 
  mutate(letter = str_sub(planet, -1,-1)) %>% 
  group_by(letter) %>% 
  summarize(n=n()) %>% 
  ggplot(aes(x = letter, y = n))+
  geom_col()

match
```




  8. The file `wordle_data.csv` contains the original solution list to the game Wordle.
In the game, players attempt to identify a secret five-letter word by guessing a word and using the color-coded information returned.
A green letter from the guess is in the correct location in the secret word.
A yellow letter from the guess is in the secret word, but is not in the correct location.
A gray letter is not in the word (unless a guess contains multiple copis of the same letter and at least one letter is colored and another is gray).

In a puzzle, the guess information is return for these first two guesses.

- SLATE returns the pattern Gray, Gray, Gray, Yellow, Gray
- COUNT returns the pattern Yellow, Green, Gray, Gray, Yellow

*(Note that the words in the data base have all lower case letters.)*

Find all possible solutions in `wrdl` which are consistent with the information given.

*Hint: you may find it easiest to pipe the word list through multiple `filter()` commands instead of trying to create a single regular expression which only matches all possible solutions. For example, you know the secret word contains at least one `T` and this letter is not in the 4th or 5th positions, at least one `C` which is not the first letter, and that the second letter is an `O`. You also know several letters which are not contained in the secret word.*
- SLATE returns the pattern Gray, Gray, Gray, Yellow, Gray
- COUNT returns the pattern Yellow, Green, Gray, Gray, Yellow

```{r}
wrdl = read_csv("../../data/wordle_data.csv") %>% 
  filter(solution == TRUE) %>%  
  select(word) %>% 
  filter(str_sub(word, 2,2 ) == "o") %>% 
  filter(str_sub(word, 1,1) != "s",
         str_sub(word, 2,2) != "s",
         str_sub(word, 3,3) != "s",
         str_sub(word, 4,4) != "s",
         str_sub(word, 5,5) != "s") %>% 
  filter(str_sub(word, 1,1) != "l",
         str_sub(word, 2,2) != "l",
         str_sub(word, 3,3) != "l",
         str_sub(word, 4,4) != "l",
         str_sub(word, 5,5) != "l") %>% 
  filter(str_sub(word, 1,1) != "a",
         str_sub(word, 2,2) != "a",
         str_sub(word, 3,3) != "a",
         str_sub(word, 4,4) != "a",
         str_sub(word, 5,5) != "a") %>% 
  filter(str_sub(word, 1,1) != "e",
         str_sub(word, 2,2) != "e",
         str_sub(word, 3,3) != "e",
         str_sub(word, 4,4) != "e",
         str_sub(word, 5,5) != "e") %>% 
  filter(str_sub(word, 1,1) != "u",
         str_sub(word, 2,2) != "u",
         str_sub(word, 3,3) != "u",
         str_sub(word, 4,4) != "u",
         str_sub(word, 5,5) != "u") %>% 
  filter(str_sub(word, 1,1) != "n",
         str_sub(word, 2,2) != "n",
         str_sub(word, 3,3) != "n",
         str_sub(word, 4,4) != "n",
         str_sub(word, 5,5) != "n") %>% 
  filter(str_sub(word, 1,1) != "c",
         str_sub(word, 2,2) == "c" |
           str_sub(word, 3,3) == "c" |
            str_sub(word, 4,4) == "c" |
              str_sub(word, 5,5) == "c") %>% 
  filter(str_sub(word, 5,5) != "t",
         str_sub(word, 4,4) != "t",
         str_sub(word, 3,3) == "t" |
           str_sub(word, 2,2) == "t" |
            str_sub(word, 1,1) == "t")
wrdl
```




